---
title: "Pulling and processing the data"
subtitle: "GM-UAW Cohort Study"
date: \today
aspectratio: 169
output:
  beamer_presentation:
    # latex_engine: lualatex
    includes:
      in_header: "~/HeadRs/BeamerHead.sty"
    keep_tex: true
    # toc: true
    # slide_level: 2
# bibliography: "`r path.expand('~/Box Sync/papers (environmental epi)/EnvEpi.bib')`"
csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
---

```{r setup, include=F}
knitr::opts_chunk$set(
	echo = F,
	warning = F,
	message = F,
	cache = F,
	fig.align = "center",
	fig.pos = "H",
	results = "asis")

library(here); library(tidyverse); library(data.table);
library(lubridate); library(xtable)

source("~/HeadRs/00-my-theme.R")
source("~/HeadRs/dot2tex.R")

```

# Overview

- Pulling and preliminary cleaning
- Building analytic dataset
	- Demographic data in person-time format
	- Map ICD codes to causes of death/cancer types
	- Merge exposure data and job history data
- Intermediate objects are saved in [Box folder 113431246688](https://berkeley.app.box.com/folder/113431246688)
- `R` code for doing all of this on [Github](https://github.com/tao-feng/gm-wrangling.git)

# Pulling and preliminary cleaning

```{r dem, eval=F}
con <- file("resources/dem.dot")
writeLines('
digraph G {

		rankdir=LR;
		node[shape=box];
		splines=ortho

		subgraph cluster_0 {

				dem[label = "\\texttt{auto\\_vs\\_15.rdata}"
						pos = "0,0!"]
				
				admin[label = "\\begin{tabular}{c}
							Trade year of hire \\\\\\\\
							for 30 pairs \\end{tabular}"
							pos = "2,0!"]
				
				unknownDeath[label = "\\begin{tabular}{c}
							10 dead with \\\\\\\\
							unknown yod \\end{tabular}"
							pos = "4,0!"]
				
				unusuYoung[label = "\\begin{tabular}{c}
							Make studyno 118137 \\\\\\\\
							5 years older \\end{tabular}"
							pos = "1,-1!"]
				
				removeImmortals[label = "\\begin{tabular}{c}
							Remove 134 people \\\\\\\\
							worked less than 3 years \\end{tabular}"
							pos = "3,-1!"]
				
				dem -> admin -> unknownDeath -> unusuYoung -> removeImmortals;
				
				label = "\\Large Demographic data";
		}
		     
 }', con)
close.connection(con)

dot2tex("dem", directory = here::here("docs/overview/resources"), prog = "fdp")
```

```{r exposure, eval=F}
con <- file("resources/exposure.dot")
writeLines('
digraph G {

		rankdir=LR;
		node[shape=box];

		subgraph cluster_1 {
				
				exposure[label = "\\begin{tabular}{c} 
						\\texttt{ganexp3.sas7bdat} (Plant 1) \\\\\\\\
						\\texttt{hanexp3.sas7bdat} (Plant 2) \\\\\\\\
						\\texttt{sanexp3.sas7bdat} (Plant 3) \\\\\\\\
						\\end{tabular}"]
						
				long[label = "\\begin{tabular}{c} 
						wide to long \\\\\\\\
						by \\texttt{studyno}, \\texttt{plant}, \\texttt{year}
						\\end{tabular}"]
						
				final[label = "\\begin{tabular}{c} 
						\\texttt{exposure.rdata} on Box \\\\\\\\
						({folder 113431246688})
						\\end{tabular}"]
						
				exposure -> long -> final
				
				label = "\\Large Annualized exposure data"
		}
		     
 }', con)
close.connection(con)

dot2tex("exposure", directory = here::here("docs/overview/resources"))

```

```{r job, eval=F}
con <- file("resources/job.dot")
writeLines('
digraph G {

		rankdir=TB;
		node[shape=box];
		splines=ortho;

		subgraph cluster_1 {
				
				job[label = "\\begin{tabular}{c} 
						\\texttt{gpp84\\_3.sas7bdat} (Plant 1) \\\\\\\\
						\\texttt{hpp84\\_3.sas7bdat} (Plant 2) \\\\\\\\
						\\texttt{spp84\\_3.sas7bdat} (Plant 3) \\\\\\\\
						\\end{tabular}"
						pos = "0,0!"]
						
				filter[label = "\\begin{tabular}{c} 
						Drop columns except for: \\\\\\\\
						\\texttt{studyno}, \\texttt{years}, \\texttt{datein}, \\texttt{dateout} \\\\\\\\
						\\texttt{histcode}, \\texttt{plant}, \\texttt{mach}, mwf components \\\\\\\\
						\\end{tabular}"
						pos = "2.55,0!"]
						
				drop[label = "\\begin{tabular}{c} 
						Drop duplicate rows
						\\end{tabular}"
						pos = "5,0!"]
						
				simplify[label = "\\begin{tabular}{c} 
						Simplify variables: \\\\\\\\
								\\begin{tabular}{rl}
								\\texttt{histcode}: & off if ``OFF\'\' or ``SR\'\';
								missing if ``MSS\'\'; \\\\\\\\
								& discrepancy if ``MD\'\' \\\\\\\\
								\\texttt{machining}: & assembly if ``AS\'\'; machining otherwise
								\\end{tabular}
						\\end{tabular}"
						pos = "0.75,-1!"]
						
				machining[label = "\\begin{tabular}{c} 
						Aggregate time spent machining/assembly \\\\\\\\
						by \\texttt{datein}/\\texttt{dateout}, \\texttt{histcode}, and \\texttt{plant}
						\\end{tabular}"
						pos = "4.5,-1!"]
						
				histcode[label = "\\begin{tabular}{c} 
						Indicator columns for \\texttt{histcode}s and mwf\\\\\\\\
						indexing by \\texttt{datein}/\\texttt{dateout} and \\texttt{plant}
						\\end{tabular}"
						pos = "0.75,-2!"]
						
				cont[label = "\\begin{tabular}{c} 
						Find gaps in \\texttt{datein}/\texttt{dateout} coverage \\\\\\\\
						(appear to be holidays; assume 0 exposure)
						\\end{tabular}"
						pos = "4,-2!"]
						
				clean[label = "\\begin{tabular}{c} 
						Convert to person-year format: \\\\\\\\
						count number of days in [\\texttt{datein}, \\texttt{dateout})
						\\end{tabular}"
						pos = "1,-3!"]
				
				final[label = "\\begin{tabular}{c} 
						\\texttt{jobhist\\_py.rdata} on Box \\\\\\\\
						({folder 113431246688})
						\\end{tabular}"
						pos = "4,-3!"]
				
				job -> filter -> drop -> simplify -> machining
				machining -> histcode -> cont -> clean -> final
				
				label = "\\Large Job history data"
		}
		     
 }', con)
close.connection(con)

dot2tex("job", directory = here::here("docs/overview/resources"), prog = "fdp")

```

\begin{minipage}{\linewidth}
	\centering
	\begin{adjustbox}{scale = 0.7}
	\input{resources/dem.tex}
	\end{adjustbox}
\end{minipage}\vspace{1em}
\begin{minipage}{\linewidth}
	\centering
	\begin{adjustbox}{scale = 0.7}
	\input{resources/exposure.tex}
	\end{adjustbox}
\end{minipage}

# 

\begin{center}
	\begin{adjustbox}{scale = 0.7}
	\input{resources/job.tex}
	\end{adjustbox}
\end{center}

# Person-year dataset

- Start with the cleaned demographic data `cohort`
- Duplicate each row so that each year from hire to death/end of FU is represented
- Index appropriately by calendar year and age

# Outcome labels

- Cause of death labels coded up using ICD mappings given by the NIOSH-92 death categories document (linked [here](https://www.cdc.gov/niosh/ltas/pdf/Rate-Info-Table-3.pdf))
- Cancer incidence from MCR coded up by Liza (thank you!)
- Cancer incidence from SEER coded up using the Site Recode ICD-O-3/WHO 2008 Definitions (linked [here](https://seer.cancer.gov/siterecode/icdo3_dwhoheme/index.html)), taking into account both the ICD-O-3 Site code _and_ the ICD-O-3 Histology code
	- Cleaned up SEER incidence data frames on Box (see `SEER incidence.csv` in (folder 113431246688)[https://berkeley.app.box.com/folder/113431246688])

# Merging exposure and job history data

- Recall that exposure data was indexed by `studyno`, `year`, _and_ `plant`
	- Before merging, exposure was summed across plants i.e. indexed by `studyno` and `year` only
- `plant` was taken to be plant with the most days in `jobhist_py.rdata`, for that `year`

# Code: it's all on github

\begin{center}
\begin{adjustbox}{max width = 0.75\linewidth}
\href{https://github.com/tao-feng/gm-wrangling.git}{\includegraphics{resources/repo.png}}
\end{adjustbox}
\end{center}

# Getting the code

```{sh, eval=F, echo=T}
#!/bin/sh 

cd "~" # Must must clone into your home directory
git clone https://github.com/tao-feng/HeadRs.git # Dependencies

cd "directory/of your/choice"
git clone https://github.com/tao-feng/gm-wrangling.git
```

Or Download ZIP -- after unzipping, please change the directory names to "HeadRs" and "gm-wrangling"

The home directory `R` sees can be found by running `path.expand("~")`

# Getting the data and helper functions

```{r, eval=F, echo=T}
# Check that necessary packages are installed
lapply(c("tidyverse", "xtable", "pander", "tikzDevice", "knitr",
	"data.table", "zoo", "boxr", "lubridate", "sas7bdat", "Hmisc"),
	function(package) {if (!package %in% installed.packages()) {
		install.packages(package)}})

# Get data and helper functions
source("directory/of your/choice/gm-wrangling/wrangling/00-hello.R")
```

Note: Setting up `boxr` is a bit of a pain...

# What you get from running `00-hello.R`

\begin{center}\begin{adjustbox}{max width = 0.85\linewidth}\ttfamily
\begin{tabular}{llll}
  \toprule
additional\_outcomes() & get.cohort\_py() & jobhist & mytheme.web \\ 
  cohort & get.exposure() & jobhist\_py & og.dir \\ 
  date.to.gm() & get.jobhist() & jobhist\_py.cast & self\_injury.function() \\ 
  death\_type() & get.ltab\_obs() & latex() & spec\_icd\_codes() \\ 
  drive\_D & gm.to.date() & ltab\_age() & tikzLualatexPackages.option \\ 
  dta & hook\_output() & ltab\_calendar() & to\_drive\_D() \\ 
  exposure & icd\_codes.function() & lualatex()  \\ 
  get.cohort\_analytic() & is.auto\_vs\_15 & mytheme  \\ 
   \bottomrule
\end{tabular}
\end{adjustbox}\end{center}

Notes

- `cohort` is the cleaned demographic data
	- Please check variable names and types
	- Most variables correspond to those in `auto_vs_15`
- `exposure` is the pulled/merged exposure data
- `jobhist_py` is the cleaned person-year job history data

# Making the analytic data object